---
title: "Phytophthora"
author: "Sam Muir"
date: "2022-12-16"
output: github_document
---

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(viridis)
library(car)
```
 
```{r echo=FALSE, include=FALSE}
phyto <- read_csv("data/plate_counts.csv")
salinity <- read_csv("soil_collection.csv")
```

## Initial data cleaning and plotting
I filtered the data to only include the counts after 4 days for both trials, as well as only including the 10-1 dilution. Then, I plotted the data to check for normal distribution. The data appears to be skewed.
```{r echo = FALSE}
phyto %>%
  filter(date_counted == "2022-11-22" | date_counted == "2022-11-14",
         dilution == "10-1") %>%
ggplot(., aes(`cfu/g`, color = location)) +
  geom_density() +
  scale_color_viridis_d() +
  theme_linedraw()
```
 
Log transforming the data shows a more normal distribution for some of them? Newtowne Neck is still skewed. 

```{r echo = FALSE}
phyto %>%
  filter(date_counted == "2022-11-22" | date_counted == "2022-11-14",
         dilution == "10-1") %>%
ggplot(., aes(log(`cfu/g`), color = location)) +
  geom_density() +
  scale_color_viridis_d() +
  theme_linedraw()
```

```{r echo=FALSE, include=FALSE}
salinity <- salinity %>%
  mutate(site = case_when(site == "chapman" ~ "CM",
                          site == "chapel" ~ "CP",
                          site == "point_lookout" ~ "PL",
                          site == "newtowne" ~ "NN")) %>%
  rename(location = site)
```

```{r echo=FALSE, include=FALSE}
phyto_log <- phyto %>%
  left_join(salinity) %>%
  rename(cfu = `cfu/g`) %>%
  mutate(trial = case_when(date_plated == "2022-11-10" ~ "A",
                           date_plated == "2022-11-18" ~ "B"),
         incubation = case_when(date_counted == "2022-11-20" ~ "2days",
                                date_counted == "2022-11-14" | date_counted == "2022-11-22" ~ "4days")) %>%
  group_by(location, dilution, trial, incubation) %>%
  filter(dilution == "10-1") %>%
  mutate(log_cfu = log(cfu),
         mean_log_cfu = mean(log_cfu),
         sd_log_cfu = sd(log_cfu))

phyto_avg <- phyto %>%
  left_join(salinity) %>%
  rename(cfu = `cfu/g`) %>%
  mutate(trial = case_when(date_plated == "2022-11-10" ~ "A",
                           date_plated == "2022-11-18" ~ "B"),
         incubation = case_when(date_counted == "2022-11-20" ~ "2days",
                                date_counted == "2022-11-14" | date_counted == "2022-11-22" ~ "4days")) %>%
  group_by(location, dilution, incubation, trial) %>%
  filter(dilution == "10-1") %>%
  mutate(mean_cfu = mean(cfu),
         sd_cfu = sd(cfu))
```

## Plotting Averages

Non-transformed data:    
- Letter indicates trial (1st or second)   
- days is incubation time
```{r}
ggplot(phyto_avg, aes(location, mean_cfu, fill = salinity_ppt)) +
  geom_bar(position="dodge", stat = "identity", size = 0.8) +
  facet_wrap(~trial + incubation) +
  geom_errorbar(aes(ymin=mean_cfu-sd_cfu, 
                    ymax=mean_cfu+sd_cfu), 
                width=0.2, linewidth = 0.5,
                 position=position_dodge(.9)) +
  theme_linedraw() +
  labs(x = "", y = "Mean cfu/g", fill = "Salinity (ppt)") +
  scale_fill_viridis(end = 0.8)
```

Log Transformed Data:    
- Letter indicates trial (1st or second)    
- days is incubation time
```{r}
ggplot(phyto_log, aes(location, mean_log_cfu, fill = salinity_ppt)) +
  geom_bar(position="dodge", stat = "identity", size = 0.8) +
  facet_wrap(~trial+incubation) +
  geom_errorbar(aes(ymin=mean_log_cfu-sd_log_cfu, 
                    ymax=mean_log_cfu+sd_log_cfu), 
                width=0.2, linewidth = 0.5,
                 position=position_dodge(.9)) +
  theme_linedraw() +
  labs(x = "", y = "Log of Mean cfu/g", fill = "Salinity (ppt)") +
  scale_fill_viridis(end = 0.8)
```

## Statistical Analysis - Log-transformed data
### For after 4 days incubation

```{r echo=FALSE, include=FALSE}
filter_aov <- phyto_log %>% 
  filter(date_counted == "2022-11-22")
```

Levene's Test
```{r}
leveneTest(log_cfu ~ location, filter_aov)
```

ANOVA
```{r}
aov <- aov(log_cfu ~ location, filter_aov)
summary(aov)
```

Tukey Test
```{r}
TukeyHSD(aov)
```

### For after 2 days incubation

```{r echo=FALSE, include=FALSE}
filter_aov2 <- phyto_log %>% 
  filter(date_counted == "2022-11-20")
```

Levene's Test
```{r}
leveneTest(log_cfu ~ location, filter_aov2)
```

ANOVA
```{r}
aov2 <- aov(log_cfu ~ location, filter_aov2)
summary(aov2)
```

Tukey Test
```{r}
TukeyHSD(aov2)
```


