---
title: "Phytophthora"
author: "Sam Muir"
date: "2022-12-16"
output: github_document
---

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(viridis)
library(car)
```
 
```{r echo=FALSE, include=FALSE}
phyto <- read_csv("data/plate_counts.csv")
salinity <- read_csv("soil_collection.csv")
```

## Initial data cleaning and plotting
I filtered the data to only include the counts after 4 days for both trials, as well as only including the 10-1 dilution. Then, I plotted the data to check for normal distribution. The data appears to be skewed.
```{r echo = FALSE}
phyto %>%
  filter(date_counted == "2022-11-22" | date_counted == "2022-11-14",
         dilution == "10-1") %>%
ggplot(., aes(`cfu/ml`, color = location)) +
  geom_density() +
  scale_color_viridis_d() +
  theme_linedraw()
```
 
Log transforming the data shows a more normal distribution for some of them? Newtowne Neck is still skewed. 

```{r echo = FALSE}
phyto %>%
  filter(date_counted == "2022-11-22" | date_counted == "2022-11-14",
         dilution == "10-1") %>%
ggplot(., aes(log(`cfu/ml`), color = location)) +
  geom_density() +
  scale_color_viridis_d() +
  theme_linedraw()
```

```{r echo=FALSE, include=FALSE}
salinity <- salinity %>%
  mutate(site = case_when(site == "chapman" ~ "CM",
                          site == "chapel" ~ "CP",
                          site == "point_lookout" ~ "PL",
                          site == "newtowne" ~ "NN")) %>%
  rename(location = site)
```

```{r echo=FALSE, include=FALSE}
phyto_salinity <- phyto %>%
  left_join(salinity) %>%
  rename(cfu = `cfu/ml`) %>%
  group_by(location, dilution) %>%
  filter(date_counted == "2022-11-22" | date_counted == "2022-11-14") %>%
  filter(dilution == "10-1") %>%
  mutate(log_cfu = log(cfu))
```


```{r finding averages, include=FALSE}
averages <- phyto_salinity %>%
  summarise(mean(log_cfu), sd(log_cfu)) %>%
  left_join(phyto_salinity)
```

## Plotting Averages
```{r}
ggplot(averages, aes(location, `mean(log_cfu)`, fill = salinity_ppt)) +
  geom_bar(position="dodge", stat = "identity", size = 0.8) +
  geom_errorbar(aes(ymin=`mean(log_cfu)`-`sd(log_cfu)`, 
                    ymax=`mean(log_cfu)`+`sd(log_cfu)`), 
                width=0.2, linewidth = 0.5,
                 position=position_dodge(.9)) +
  theme_linedraw() +
  scale_x_discrete(labels=c("CM" = "Chapman", "CP" = "Chapel Point",
                              "NN" = "Newtowne Neck", "PL" = "Point Lookout")) +
  labs(x = "", y = "Log of Mean cfu/mL", fill = "Salinity (ppt)") +
  scale_fill_viridis(end = 0.8)
```

## Statistical Analysis
Levene's Test
```{r}
leveneTest(log_cfu ~ location, averages)
```
Since p > 0.05 there is equal variance and we can perform an ANOVA.

ANOVA
```{r}
aov <- aov(log_cfu ~ location, averages)
summary(aov)
```

